### 2013-08-26 doc/dissertation/document/chapters/introduction/introduction.tex
adding pauls comments through chapter 3

To achieve these goals, three complementary research and development
activities are performed. First, Monte Carlo Sythetic Acceleration
(MCSA) is applied to the neutron transport problem in a light water
reactor calculation to develop the necessary preconditioning
techniques and understanding required to solve difficult
problems. Second, using these new techniques, a series of benchmarks
for the Navier-Stokes equations are solved in different flow regimes
to research the performance of the Monte Carlo algorithm for nonlinear
problems using the new Forward-Automated Newton-MCSA (FANM) algorithm
developed by this work. Third, a novel parallel algorithm based on
domain decomposed Monte Carlo methods for particle transport is
developed to fully parallelize the new solution techniques on
leadership class hardware with scaling studies performed in this work
using a neutron diffusion problem.

>>>>

To achieve these goals, three complementary research and development
activities are performed. First, Monte Carlo Synthetic Acceleration
(MCSA) is applied to the neutron transport problem in a light water
reactor calculation to develop the necessary preconditioning
techniques and understanding required to solve difficult
problems. Second, using these new techniques, a series of benchmarks
for the Navier-Stokes equations are solved in different flow regimes
to research the performance of the Monte Carlo algorithm for nonlinear
problems using the new Forward-Automated Newton-MCSA (FANM) algorithm
developed by this work. Third, a novel parallel algorithm based on
domain decomposed Monte Carlo methods for particle transport is
developed to fully parallelize the new solution techniques on
leadership class hardware with scaling studies performed in this work
using a neutron diffusion problem.

### 2013-08-26 doc/dissertation/document/chapters/spn_equations/spn_equations.tex
adding pauls comments through chapter 3

%%---------------------------------------------------------------------------%%
\section{Advanced Preconditioning Strategies}
\label{subsec:spn_advanced_preconditioning}
For the fuel assembly criticality problem presented in the previous
section, the spectral radius was near one using Jacobi preconditioning
and in a regime in which MCSA breakdown is observed. In this regime
the number of stochastic histories required to converge MCSA increases
rapidly and the resulting poor performance is compounded by those
histories being increasingly expensive to compute. To overcome this
difficulty, advanced preconditioning strategies for the $SP_N$ problem
are required beyond simple Jacobi methods that can reduce the spectral
radius into a region of better MCSA performance. Two modern algebraic
preconditioning strategies, ILUT and SPAINV, will be presented here
and used within the explicit preconditioning framework given by
Eq~(\ref{eq:left_right_mcsa}). Data showing their effects on MCSA
solutions of the fuel assembly criticality problem will be provided.

\subsection{ILUT Preconditioning}
\label{subsec:spn_ilut_preconditioning}
ILUT preconditioning is chosen first due to its applicability to
problems that have a dominating elliptic character such as the fuel
assembly problem or other problems with a large scattering to
absorption ratio. The algebraic formulation of ILUT requires the
physics operator to be fully formed, also a requirement for MCSA, and
is therefore easily incorporated into our solution sequence.

Incomplete lower-upper (ILU) factorizations of the linear operator can
be used as a simple mechanism to form an approximate inverse of a
preconditioner. To build the factorization, the sparse upper and lower
triangular factors, $\mathbf{L}$ and $\mathbf{U}$, are computed such
that the residual matrix formed by the factorization
\cite{saad_iterative_2003}:
\begin{equation}
  \mathbf{R} = \mathbf{L} \mathbf{U} - \mathbf{A} \:,
  \label{eq:ilu_residual_matrix}
\end{equation}
has a specified sparsity pattern and element magnitude. When a
magnitude threshold is used, elements generated in the factors below
that magnitude are dropped, resulting in ILU threshold (ILUT)
preconditioning. The sparsity pattern in this case is determined from
the input matrix to be preconditioned and the number of elements
maintained in the factor is specified by a fill level parameter. A
fill level of 1 will generate the same number of elements as the
sparsity pattern of the input matrix while a fill level of 2 will
contain twice as many elements in the factorization, resulting in a
better representation of the true LU factorization. The inverse of the
lower and upper triangular factors may then be easily inverted by
means of simple elimination procedure to produce the
preconditioner. For MCSA, $\mathbf{L}^{-1}$ will be used on the left
and $\mathbf{U}^{-1}$ on the right to precondition the system.

For the explicit preconditioning scheme presented in
Eq~(\ref{eq:left_right_mcsa}), the fully inverted operator must be
generated in order to build the set of probabilities and weights
required for Monte Carlo sampling. Therefore, for a linear system of
size $N$, $N$ triangular solves will be required in order to extract
the inverse matrices from production ILU implementations. In addition,
parallel implementations typically generate lower and upper factors
that are only triangular locally, providing an easily parallelizable
mechanism to generate the action of the preconditioner inverse. A
consequence of this choice for parallel scalability is a degradation
of the preconditioning quality as the size of the parallel system is
increased. For serial computations, the triangular factorization is
potentially exact depending on the parameters chosen while at
thousands of parallel tasks, the global triangular factors differ
significantly from the true factorization. As a result, more
iterations are required at higher levels of parallelism to converge
the system and will ultimately degrade overall scalability of the
system with respect to total wall time to converge.

MCSA preconditioned with ILUT was used to solve the fuel assembly
problem presented in the previous section for a 1-group $SP_1$
discretization. Unlike the Jacobi preconditioned strategy, convergence
was achieved with ILUT preconditioning. To study convergence
sensitivity to ILUT parameters, the fill level and drop tolerance were
parametrically varied with the number of iterations to converge an
eigenvalue iteration for the fuel assembly problem recorded along with
the maximum number of non-zero entries observed in all matrix rows for
the left/right preconditioned composite linear operator. To provide
some sparsity to the factorization, the ILUT drop tolerance was used
to drop elements in the extracted inverse triangular factors. For each
calculation, the number of iterations required to converge reported
was for a single eigenvalue iteration with \sn{3}{4} histories at each
MCSA iteration to compute the Monte Carlo correction using the adjoint
collision estimator. All ILUT calculations reported here were
performed on a single CPU and therefore this data does not take into
account the aforementioned effects of parallel decomposition on the
quality of the ILUT preconditioning.

@@

As a means to assess the quality of the preconditioning, a simple
metric is developed to address these concerns and allow comparison to
future developments. For our studies, our core performance metric will
be iterative performance with the minimum number of MCSA iterations
required to converge the problem desired. For Monte Carlo
calculations, this improved performance is balanced by the creation of
a dense composite system and we seek to reduce the number of non-zero
entries to a minimum value in order to lower the amount of coupling
among states in the system and reduce the amount of memory used
along with the potential latency overhead. Given these two objectives,
the following metric is proposed:
\begin{equation}
  \text{Quality Metric} = (\text{\# iterations}) \times (\text{maximum
    \# non-zero values})\:,
\end{equation}
where the highest quality preconditioning is one that minimizes this
metric. For the ILUT preconditioned data provided,
Figure~\ref{fig:ilut_quality} provides the computed metric as a
function of ILUT drop tolerance for each fill level used. In general,
the metric is similar to the data observed for the number of
iterations required to converge as the non-zero row entries changed by
a smaller fraction over drop tolerances tested. For higher levels of
fill, the quality metric is observed to be relatively insensitive to
the drop tolerance as maintaining more entries in the preconditioner
has a stronger impact than filtering small values.

@@

\subsection{Sparse Approximate Inverse Preconditioning}
\label{subsec:spn_spainv_preconditioning}
Sparse approximate inverse (SPAINV) preconditioning is chosen next because the true inverse of the preconditioner
is product of the algorithm, avoiding the need for many inverse matrix
applications as required by using ILUT with MCSA. In addition, the
algorithm seeks to maintain the sparsity of the inverse
preconditioner, potentially alleviating some of the dense operator
issues observed with ILUT.

@@

Figure~\ref{fig:spainv_iterations} gives the number of MCSA iterations
needed to converge the fuel assembly problem and
Figure~\ref{fig:spainv_size} the maximum number of non-zero entries
per row observed in the composite linear operator as a function the
SPAINV threshold. For this analysis, the number of levels in the
sparsity pattern was varied from 3 to 7 with convergence of the fuel
assembly problem not achieved for smaller level values. Compared to
ILUT preconditioning, at higher levels we see comparable iterative
performance with a relative invariance to the threshold value. The
threshold did have a significant effect, however, on the time required
to generate the preconditioner with a threshold of 0.001 over an order
of magnitude slower than a threshold of 0.01 due to the inclusion of a
significant number of extra values in the input operator. In addition
to comparable iterative performance, the sparsity of the composite
operator is greatly improved with nearly an order of magnitude
reduction in number of non-zero entries per row for lower level
values.

@@

For each row in the preconditioned system, we can then apply the
reduced domain approximation to eliminate those elements that either
fall below a specified tolerance level, keep the $N$ largest elements
in each row, or both giving:
\begin{subequations}
  \begin{gather}
    \ve{x}^{k+1/2} = \ve{x}^k + \ve{r}^k\:,\\
    \ve{r}^{k+1/2} = \ve{b} - \ve{A}\ve{x}^{k+1/2}\:,\\
    \hat{\ve{A}}\delta\ve{x}^{k+1/2} = \ve{r}^{k+1/2}\:, 
    \label{eq:rda_correction}\\
    \ve{x}^{k+1} = \ve{x}^{k+1/2} + \delta \ve{x}^{k+1/2}\:,\\
    \ve{r}^{k+1} = \ve{b} - \ve{A}\ve{x}^{k+1}\:,
  \end{gather}
  \label{eq:rda_mcsa}
\end{subequations}
where $\hat{\ve{A}}$ in Eq~(\ref{eq:rda_correction}) means that the
reduced domain approximation has been applied to the iteration matrix
in the Monte Carlo method. If enough of the terms in the original
iteration matrix have been maintained, the character of the original
iteration matrix should be retained and continue to accelerate the
iterative scheme.

@@

Table~\ref{tab:spn_comparison_iterations} gives the average number of
linear solver iterations required to converge a single eigenvalue
iteration of the fuel assembly problem as a function of the number of
energy groups. In general, the iterative performance of all the
solvers tested aside from the Richardson iteration is comparable with
BiCGStab performing the best. We expect this because not only are the
multigroup $SP_N$ equations positive-definite, but they are also
block-symmetric and therefore we expect the conjugate gradient-based
methods to perform well due to the structure of the projection
scheme. It important to note that the iterative performance of each
solver is not a strong function of the number of energy groups in the
problem (and therefore the problem size). Even more important,
although BiCGStab produced the best iterative results, MCSA did
perform better than GMRES in many cases with the Richardson iteration
clearly accelerated by the Monte Carlo solve. When compared to the
preconditioned Richardson iteration, the acceleration provided by MCSA
is clear with 3-4 times fewer iterations required for convergence.

@@

Although iterative performance for MCSA was comparable to that
observed for Krylov implementations using exactly the same
preconditioning scheme, timing performance is not as
competitive. Figure~\ref{fig:spn_comparison_time} gives the average
CPU time per linear solver iteration required to converge the fuel
assembly problem with all linear solver iterations over all eigenvalue
iterations used to compute the average. The Krylov solver
implementations are approximately an order of magnitude faster than
the MCSA implementations presented here. We expect these results for
several reasons. First, even when the reduced domain approximation is
used there are $O(100)$ elements in each row of the system and each of
those elements must be processed during a transition in the Monte
Carlo random walk sequence. In addition, although the preconditioned
fixed point iterations used in the MCSA sequence do not use the
composite linear operator but rather a sequence of matrix vector
multiplies to achieve the same preconditioning effect, they do use the
explicitly extracted inverse of the preconditioning matrices which
themselves are dense, leading to exceedingly slow computation times
even in the fixed point iteration.

@@

Based on the performance results in this section, MCSA shows promise
as a competitive and perhaps even superior method for solutions to the
neutron transport problem discretized with the $SP_N$
approximation. Not only are the correct answers produced when compared
to production linear solvers, but the iterative performance is
comparable to Krylov methods with identical preconditioning that would
typically be used every-day calculations. From a CPU timing
perspective, the methods yield the same time complexity as a function
of the number of energy groups in the problem when compared to the
Krylov methods. Here, a large time constant is generated due to the
explicit preconditioning strategy and the generation of dense linear
operators as a result. To improve these results and put general MCSA
schemes into a performance regime where they are competitive with
Krylov methods for neutron transport problems, significant research
will be required to improve upon the explicit preconditioning scheme
presented here.

@@

\begin{itemize}
\item MCSA can solve the asymmetric system generated by the $SP_N$
  equations
\item MCSA has been incorporated into the Exnihilo neutronics
  production code base developed at Oak Ridge National Laboratory
\item Light water reactor problems are difficult to solve with MCSA as
  they have large spectral radii due to the neutron scattering in the
  moderator
\item Simple Jacobi-based preconditioning reduces the spectral radius
  of the $SP_N$ system below unity, however, this alone is not
  sufficient for MCSA convergence
\item As the spectral radius approaches unity, MCSA was demonstrated
  to break down with more stochastic histories required for
  convergence and more CPU time required per history
\item Advanced algebraic preconditioning strategies were applied to
  the $SP_N$ equations to obtain convergence with ILUT chosen for
  subsequent investigations
\item The reduced domain approximation was introduced to mitigate the
  memory and performance constraints of the explicit preconditioning
  strategy although both remain significant obstacles
\item MCSA was verified to produce the same k-eigenvalue for the fuel
  assembly as production Krylov methods and gave good relative flux
  agreement
\item MCSA was observed to converge in fewer iterations per eigenvalue
  iteration than GMRES for the fuel assembly criticality problem and
  more than Bi-CGStab using the same preconditioning
\item The explicit preconditioning strategy required to overcome the
  MCSA spectral radius restriction creates runtimes $O(100)$ larger
  than the production Krylov methods for the fuel assembly criticality
  problem
\end{itemize}

>>>>

%%---------------------------------------------------------------------------%%
\section{Advanced Preconditioning Strategies}
\label{subsec:spn_advanced_preconditioning}
For the fuel assembly criticality problem presented in the previous
section, the spectral radius was near one using Jacobi preconditioning
and in a regime in which MCSA breakdown is observed. In this regime
the number of stochastic histories required to converge MCSA increases
rapidly and the resulting poor performance is compounded by those
histories being increasingly expensive to compute. To overcome this
difficulty, advanced preconditioning strategies for the $SP_N$ problem
are required, beyond simple Jacobi methods, that can reduce the
spectral radius into a region of better MCSA performance. Two modern
algebraic preconditioning strategies, ILUT and SPAINV, will be
presented here and used within the explicit preconditioning framework
given by Eq~(\ref{eq:left_right_mcsa}). Data showing their effects on
MCSA solutions of the fuel assembly criticality problem will be
provided.

\subsection{ILUT Preconditioning}
\label{subsec:spn_ilut_preconditioning}
ILUT preconditioning is chosen first due to its applicability to
problems that have a dominating elliptic character such as the fuel
assembly problem or other problems with a large scattering to
absorption ratio \cite{saad_iterative_2003}. The algebraic formulation
of ILUT requires the physics operator to be fully formed, also a
requirement for MCSA, and is therefore easily incorporated into our
solution sequence.

Incomplete lower-upper (ILU) factorizations of the linear operator can
be used as a simple mechanism to form an approximate inverse of a
preconditioner. To build the factorization, the sparse upper and lower
triangular factors, $\mathbf{L}$ and $\mathbf{U}$, are computed such
that the residual matrix formed by the factorization
\cite{saad_iterative_2003}:
\begin{equation}
  \mathbf{R} = \mathbf{L} \mathbf{U} - \mathbf{A} \:,
  \label{eq:ilu_residual_matrix}
\end{equation}
has a specified sparsity pattern and element magnitude. When a drop
tolerance threshold is used, elements generated in the factors below
that magnitude are dropped, resulting in ILU threshold (ILUT)
preconditioning. The sparsity pattern in this case is determined from
the input matrix to be preconditioned and the number of elements
maintained in the factor is specified by a fill level parameter. A
fill level of 1 will generate the same number of elements as the
sparsity pattern of the input matrix while a fill level of 2 will
contain twice as many elements in the factorization, resulting in a
better representation of the true LU factorization. The inverse of the
lower and upper triangular factors may then be easily inverted by
means of simple elimination procedure to produce the
preconditioner. For MCSA, $\mathbf{L}^{-1}$ will be used on the left
and $\mathbf{U}^{-1}$ on the right to precondition the system.

For the explicit preconditioning scheme presented in
Eq~(\ref{eq:left_right_mcsa}), the fully inverted preconditioners must
be generated in order to build the set of probabilities and weights
required for Monte Carlo sampling. Therefore, for a linear system of
size $N$, $N$ triangular solves with the basis of each row in the
system as a solution vector will be required in order to extract the
inverse matrices from production ILU implementations. In addition,
parallel implementations typically generate lower and upper factors
that are only triangular locally, providing an easily parallelizable
mechanism to generate the action of the preconditioner inverse. A
consequence of this choice for parallel scalability is a degradation
of the preconditioning quality as the size of the parallel system is
increased. For serial computations, the triangular factorization is
potentially exact depending on the parameters chosen while at
thousands of parallel tasks, the global triangular factors differ
significantly from the true factorization. As a result, more
iterations are required at higher levels of parallelism to converge
the system and will ultimately degrade overall scalability of the
system with respect to total wall time to converge.

MCSA preconditioned with ILUT was used to solve the fuel assembly
problem presented in the previous section for a 1-group $SP_1$
discretization. Unlike the Jacobi preconditioned strategy, convergence
was achieved with ILUT preconditioning. To study convergence
sensitivity to ILUT parameters, the fill level and drop tolerance were
parametrically varied with the number of iterations to converge an
eigenvalue iteration for the fuel assembly problem recorded. In
addition, the maximum number of non-zero entries observed in all rows
for the left/right preconditioned composite linear operator was
recorded to observe the effects of the explicit preconditioner
inversion and matrix-matrix multiplication on sparsity. To provide
some sparsity to the factorization, an additional step was added with
the ILUT drop tolerance used to drop elements in the explicitly
inverted preconditioners. For each calculation, the number of
iterations reported was for a single eigenvalue iteration with
\sn{3}{4} histories at each MCSA iteration to compute the Monte Carlo
correction using the adjoint collision estimator. All ILUT
calculations reported here were performed on a single CPU and
therefore this data does not take into account the aforementioned
effects of parallel decomposition on the quality of the ILUT
preconditioning.

@@

As a means to assess the quality of the preconditioning, a simple
metric is developed to address these concerns and allow comparison to
future developments. For our studies, our core performance metric will
be iterative performance with the minimum number of MCSA iterations
required to converge the problem desired. For Monte Carlo
calculations, this improved performance is balanced by the creation of
a dense composite system and we seek to reduce the number of non-zero
entries to a minimum value in order to lower the amount of coupling
among states in the system and reduce the amount of memory used along
with the potential latency overhead. Given these two objectives, the
following metric is proposed:
\begin{equation}
  \text{Quality Metric} = (\text{\# iterations}) \times (\text{maximum
    \# non-zero values})\:,
\end{equation}
where the highest quality preconditioning is one that minimizes this
metric. This metric gives equal importance to both iterative
performance and sparsity with the purpose of simply measuring their
combined effect rather than seeking some optimal relationship of
importance. For the ILUT preconditioned data provided,
Figure~\ref{fig:ilut_quality} provides the computed metric as a
function of ILUT drop tolerance for each fill level used. In general,
the metric is similar to the data observed for the number of
iterations required to converge as the non-zero row entries changed by
a smaller fraction over drop tolerances tested. For higher levels of
fill, the quality metric is observed to be relatively insensitive to
the drop tolerance as maintaining more entries in the preconditioner
has a stronger impact than filtering small values.

@@

\subsection{Sparse Approximate Inverse Preconditioning}
\label{subsec:spn_spainv_preconditioning}
Sparse approximate inverse (SPAINV) preconditioning is chosen next
because the true inverse of the preconditioner is the product of the
algorithm, avoiding the need for many inverse matrix applications as
required by using ILUT with MCSA. In addition, the algorithm seeks to
maintain the sparsity of the inverse preconditioner, potentially
alleviating some of the dense operator issues observed with ILUT.

@@

Figure~\ref{fig:spainv_iterations} gives the number of MCSA iterations
needed to converge the fuel assembly problem and
Figure~\ref{fig:spainv_size} the maximum number of non-zero entries
per row observed in the composite linear operator as a function the
SPAINV threshold. For this analysis, the number of levels in the
sparsity pattern was varied from 3 to 7 with convergence of the fuel
assembly problem not achieved for smaller level values. Compared to
ILUT preconditioning, at higher levels we see comparable iterative
performance with a relative invariance to the threshold value. The
threshold did have a significant effect, however, on the time required
to generate the preconditioner with a threshold of 0.001 over an order
of magnitude slower than a threshold of 0.01 due to the inclusion of a
significant number of extra values in the input operator,
$\mathbf{A}$. In addition to comparable iterative performance, the
sparsity of the composite operator is greatly improved with nearly an
order of magnitude reduction in number of non-zero entries per row for
lower level values.

@@

For each row in the preconditioned system, we can then apply the
reduced domain approximation to eliminate those elements that either
fall below a specified threshold level, fall within a fill level $N$
where we keep the $N$ largest elements in each row, or both giving:
\begin{subequations}
  \begin{gather}
    \ve{x}^{k+1/2} = \ve{x}^k + \ve{r}^k\:,\\
    \ve{r}^{k+1/2} = \ve{b} - \ve{A}\ve{x}^{k+1/2}\:,\\
    \hat{\ve{A}}\delta\ve{x}^{k+1/2} = \ve{r}^{k+1/2}\:, 
    \label{eq:rda_correction}\\
    \ve{x}^{k+1} = \ve{x}^{k+1/2} + \delta \ve{x}^{k+1/2}\:,\\
    \ve{r}^{k+1} = \ve{b} - \ve{A}\ve{x}^{k+1}\:,
  \end{gather}
  \label{eq:rda_mcsa}
\end{subequations}
where $\hat{\ve{A}}$ in Eq~(\ref{eq:rda_correction}) means that the
reduced domain approximation has been applied to the iteration matrix
in the Monte Carlo method. If enough of the terms in the original
iteration matrix have been maintained, the character of the original
iteration matrix should be retained and continue to accelerate the
iterative scheme.

@@

Table~\ref{tab:spn_comparison_iterations} gives this average number of
linear solver iterations required to converge a single eigenvalue
iteration of the fuel assembly problem as a function of the number of
energy groups. In general, the iterative performance of all the
solvers tested aside from the Richardson iteration is comparable with
BiCGStab performing the best. We expect this because not only are the
multigroup $SP_N$ equations positive-definite, but they are also
block-symmetric and therefore we expect the conjugate gradient-based
methods to perform well due to the structure of the projection
scheme. It is important to note that the iterative performance of each
solver is not a strong function of the number of energy groups in the
problem (and therefore the problem size). Even more important,
although BiCGStab produced the best iterative results, MCSA did
perform better than GMRES in many cases with the Richardson iteration
clearly accelerated by the Monte Carlo solve. When compared to the
preconditioned Richardson iteration, the acceleration provided by MCSA
is clear with 3-4 times fewer iterations required for convergence.

@@

Although iterative performance for MCSA was comparable to that
observed for Krylov implementations using exactly the same
preconditioning scheme, timing performance is not as
competitive. Figure~\ref{fig:spn_comparison_time} gives the average
CPU time per linear solver iteration required to converge the fuel
assembly problem with all linear solver iterations over all eigenvalue
iterations used to compute the average. The Krylov solver
implementations are approximately an order of magnitude faster than
the MCSA implementations presented here. We expect these results for
several reasons. First, even when the reduced domain approximation is
used there are $O(100)$ elements in each row of the system and each of
those elements must be accessed for next-event sampling and tally
vector modification during a transition in the Monte Carlo random walk
sequence. In addition, although the preconditioned fixed point
iterations used in the MCSA sequence do not use the composite linear
operator but rather a sequence of matrix vector multiplies to achieve
the same preconditioning effect, they do use the explicitly extracted
inverse of the preconditioning matrices which themselves are dense,
leading to exceedingly slow computation times even in the fixed point
iteration.

@@

Based on the performance results in this section, MCSA shows promise
as a competitive method for solutions to the neutron transport problem
discretized with the $SP_N$ approximation and could potentially be
superior if the preconditioning strategy were improved to reduce time
to solution and memory usage. Not only are the correct answers
produced when compared to production linear solvers, but the iterative
performance is comparable to Krylov methods with identical
preconditioning that would typically be used most calculations. From a
CPU timing perspective, the methods yield the same time complexity as
a function of the number of energy groups in the problem when compared
to the Krylov methods. Here, a large time constant is generated due to
the explicit preconditioning strategy and the generation of dense
linear operators as a result. To improve these results and put general
MCSA schemes into a performance regime where they are competitive with
Krylov methods for neutron transport problems, significant research
will be required to improve upon the explicit preconditioning scheme
presented here.

@@

\begin{itemize}
\item MCSA can solve the asymmetric system generated by the $SP_N$
  equations
\item MCSA has been incorporated into the Exnihilo neutronics
  production code base developed at Oak Ridge National Laboratory
\item Light water reactor problems are difficult to solve with MCSA as
  they have large spectral radii due to the neutron scattering in the
  moderator
\item Simple Jacobi-based preconditioning reduces the spectral radius
  of the $SP_N$ system below unity, however, this alone is not
  sufficient for MCSA convergence
\item As the spectral radius approaches unity, MCSA was demonstrated
  to break down with more stochastic histories required for
  convergence and more CPU time required per history
\item Advanced algebraic preconditioning strategies were applied to
  the $SP_N$ equations to obtain convergence with ILUT chosen for
  subsequent investigations
\item The reduced domain approximation was introduced to mitigate the
  memory and performance constraints of the explicit preconditioning
  strategy although both remain significant obstacles
\item MCSA was verified to produce the same k-eigenvalue for the fuel
  assembly as production Krylov methods and gave good relative flux
  agreement
\item MCSA was observed to converge in fewer iterations per eigenvalue
  iteration than GMRES for the fuel assembly criticality problem and
  more than Bi-CGStab using the same preconditioning
\item The explicit preconditioning strategy required to overcome the
  MCSA spectral radius restriction creates run times $O(100)$ larger
  than the production Krylov methods for the fuel assembly criticality
  problem
\end{itemize}

### 2013-08-26 doc/dissertation/document/frontmatter/abstract.tex
adding pauls comments through chapter 3

Third, a new domain decomposed algorithm to parallelize MCSA aimed at
leveraging leadership-class computing facilities was developed by
utilizing parallel strategies from the radiation transport
community. The new algorithm utilizes the Multiple-Set
Overlapping-Domain strategy in an attempt to reduce parallel overhead
and add a natural element of replication to the algorithm. A large
number of parallel scaling studies using a neutron diffusion problem
were performed on the Titan Cray XK7 machine at Oak Ridge National
Laboratory with up to 65,356 cores used in the strong scaling studies
and 131,072 cores used in the weak scaling studies. It was found that
for the current implementation of MCSA, both weak and strong scaling
improved on that observed for production implementations of Krylov
methods. In addition, the replication in the algorithm was able to
boost parallel efficiencies by up to 20\% at large core counts in the
strong scaling studies and enhance time to solution in the weak
scaling stduies by using the principle of superposition. Finally, it
was discovered that MCSA exhibits the most scalable behavior when it
is formulated as a stochastic realization of an additive Schwarz
method.

>>>>

Third, a new domain decomposed algorithm to parallelize MCSA aimed at
leveraging leadership-class computing facilities was developed by
utilizing parallel strategies from the radiation transport
community. The new algorithm utilizes the Multiple-Set
Overlapping-Domain strategy in an attempt to reduce parallel overhead
and add a natural element of replication to the algorithm. A large
number of parallel scaling studies using a neutron diffusion problem
were performed on the Titan Cray XK7 machine at Oak Ridge National
Laboratory with up to 65,356 cores used in the strong scaling studies
and 131,072 cores used in the weak scaling studies. It was found that
for the current implementation of MCSA, both weak and strong scaling
improved on that observed for production implementations of Krylov
methods. In addition, the replication in the algorithm was able to
boost parallel efficiencies by up to 20\% at large core counts in the
strong scaling studies and enhance time to solution in the weak
scaling studies by using the principle of superposition. Finally, it
was discovered that MCSA exhibits the most scalable behavior when it
is formulated as a stochastic realization of an additive Schwarz
method.
